SPARQL on the Open, Decentralised Web

### 2016-06-23 dissertation/introduction.tex
a bit on SPARQL

@@

@@

>>>>

\subsection{Basic technology}

@@

Note that all the IRIs uses the HTTP schema. The Hypertext Transfer
Protocol (HTTP) is the main application protocol used on the Internet
for the Web, and is also the protocol used most for the Semantic
Web. It is a request-response protocol in a client-server computing
model, but also accommodates for intermediate proxies.

@@

\subsection{Interacting With Data}

To read and write this graph, several techniques have been
developed. For reading, one is to simply parse the string into a
suited data structure and use any tool available in the chosen
programming language to examine the data structure, or simply write
data to e.g. a Turtle string.

A more sophisticated approach is to match parts of the graphs using
\emph{triple patterns}. A triple pattern may contain variables for
certain terms, for example:

\begin{verbatim}
?subject a ?class .
\end{verbatim}
This triple pattern will match the two triples 
\begin{verbatim}
<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document .
\end{verbatim}
from the above example. Frameworks in popular programming languages
commonly have a method that implements this.

The triple pattern is also a fundamental building block of the SPARQL
query language, which is the language that is the main topic of this
dissertation. SPARQL is an extensive, standardised query language with
both read and write operations. The syntax is near Turtle, but as the
above triple patterns, introduces variables. A set of triple patterns
may be used to create a conjunctive query, and along with a filter
that can be used to further constrain the query with boolean
expressions. Together, they are known as a Basic Graph Pattern
(BGP). More advanced parts of the language is built around BGPs.

SPARQL also introduced a mechanism for naming a graph or parts of a
graph with a IRI, an approach that was subsequently adopted by
RDF. Today, RDF is not just a triple, but may be a quad.
